### 2024-06-20 Simulation1.Rmd
A first serious commit

---
title: "Simulation"
output: pdf_document
date: "2024-06-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
```

```{r}
trees <- read_excel("~/work/Sim_article2024/artificial_forest_round.xls")
```

```{r}
# First visualisation
plot(x = trees$x, y = trees$y, main = "Forest")
```

```{r}
# True values
cat("Il y a au total", length(trees$x), "arbres et un volume total de", sum(trees$v), "\n", "Soit un volume moyen de", mean(trees$v), "\n", "Le diamètre moyen à hauteur de poitrine est de", mean(trees$d130))
```

```{r}
plot(x = trees$x, y = trees$y, main = "Forest", panel.first=grid())

```

```{r}
# Create another variable to indicate in which cell is each tree
for (i in vec){

}
```

```{r}
list <- sample(1:10)
```

```{r}
n <- 600 # dimension of a cell = 1 unit x 1 unit
vec <- c() # vector containing coordinates of cells
for (j in 0:9){ # for each number (1 to 10)
  for (i in 0:((n/10)-1)){ # translation
    vec <- append(vec, 200+j+10*i)
  }
}
```

```{r}
trees$cell2 <- 1*as.numeric((trees$x %in% seq(200, 300, by = .01)) & (trees$y %in% seq(200, 300, by = .01))) # Problem with 107th tree e.g.
trees$cell <- 1*as.numeric((200 <= trees$x) & (trees$x < 300) & (200 <= trees$y) & (trees$y < 300))
# trees[trees$cell == 1 & trees$cell2 == 0,]
# Where is the pb?
# 297.21 %in% seq(233, 300, by = .01) # FALSE, idem for numbers below 233
# 297.21 %in% seq(234, 300, by = .01) # TRUE, idem for numbers above 234
```

```{r}
for(l in vec){

}
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```