### 2024-06-24 Draft.Rmd
Excluding some draft

---
title: "Draft"
output: html_document
date: "2024-06-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Draft

```{r}
# Tests (preliminaries)

# trees$cell2 <- 1*as.numeric((trees$x %in% seq(200, 300, by = .01)) & (trees$y %in% seq(200, 300, by = .01))) # Problem with 107th tree e.g.
# trees$cell <- 1*as.numeric((200 <= trees$x) & (trees$x < 300) & (200 <= trees$y) & (trees$y < 300))

# trees[trees$cell == 1 & trees$cell2 == 0,]
# Where is the pb?
# 297.21 %in% seq(233, 300, by = .01) # FALSE, idem for numbers below 233
# 297.21 %in% seq(234, 300, by = .01) # TRUE, idem for numbers above 234
```

```{r}
# Un peu long

# PROBLEME: ne garde que le 8 (i.e. le dernier nb)

trees$cell <- rep(0, 30942) # Initialisation de la variable

for (nb in 1:10){ # All the categories (to identify the cells)
for (ligne in 0:599){ # All the lines (en parcourant par rapport aux ordonnées)
  for(c in 1:60){ # En parcourant sur une même ligne
    # if (trees$cell == 0)
    trees$cell <- ifelse(
    # (trees$cell == 0) & 
      (vec[(nb-1)*60 + 1] <= trees$x) & (trees$x < vec[(nb-1)*60 + 1]+1) & (200+ligne <= trees$y) & (trees$y < 200+ligne+1), 
    list[nb], trees$cell)
  }

}
}

# (ligne-1)*trend + 
# (ligne-1)*trend + 
```

```{r}
# Un peu long

# PROBLEME: ne garde que le 8 (i.e. le dernier nb)

trees$cell <- rep(0, 30942) # Initialisation de la variable

for (ligne in 0:599){ # All the lines (en parcourant par rapport aux ordonnées)
  # for(c in 1:60){ # En parcourant sur une même ligne
    for (nb in 1:10){ # All the categories (to identify the cells)
    # if (trees$cell == 0)
    trees$cell <- ifelse(
    # (trees$cell == 0) & 
      (vec[(nb-1)*60 + 1] <= trees$x) & (trees$x < vec[(nb-1)*60 + 1]+1) & (200+ligne <= trees$y) & (trees$y < 200+ligne+1), 
    list[nb], trees$cell)
  }

# }
}
```

```{r}
unique(trees$cell)
```

```{r}
trees$cell3 <- rep(0, 30942)
trees$cell3 <- ifelse(
  (vec[1] <= trees$x) & (trees$x < vec[1]+1) & (vec[1] <= trees$y) & (trees$y < vec[1]+1), 
  1, trees$cell3)
```

```{r}
# Inaccurate cell
trees$cell3 <- ifelse(
  (200 <= trees$x) & (trees$x < 300) & (200 <= trees$y) & (trees$y < 300), 
  1, trees$cell3)
```

```{r}
# Invalid
trees$cell3 <- rep(0, 30942)
for (ligne in 1:60){

  for(l in 1:60){
    trees$cell3 <- ifelse(
    (vec[l] <= trees$x) & (trees$x < vec[l]+1) & (vec[1] <= trees$y) & (trees$y < vec[1]+1), 
    list[1], trees$cell3)
  }

}
```

```{r}
# When only focusing on 1 10th of the population
plot(x = trees[trees$cell3 == 9,]$x, y = trees[trees$cell3 == 9,]$y, main = "Forest", panel.first=grid())
# Pretty nice (at one point)
```

```{r}
# Test

trees$cell3 <- rep(0, 30942)
nb = 1
ligne = 0

for(c in 1:60){ # En parcourant sur une même ligne
    # if (trees$cell == 0)
    trees$cell <- ifelse(
    (vec[(c-1)*trend + (nb-1) + 1] <= trees$x) & (trees$x < vec[(c-1)*trend + nb]+1) & (200+ligne <= trees$y) & (trees$y < 200+ligne+1) & trees$cell, 
    list[nb], trees$cell)
  }
```

```{r}
# When only focusing on 1 10th of the population
plot(x = trees[trees$cell == 1,]$x, y = trees[trees$cell == 1,]$y, main = "Forest", panel.first=grid())
```

```{r}
# When only focusing on 
plot(x = trees[trees$cell == 9,]$x, y = trees[trees$cell == 9,]$y, main = "Forest", panel.first=grid())
```

```{r}
# When only focusing on 
plot(x = trees[trees$cell == 0,]$x, y = trees[trees$cell == 0,]$y, main = "Forest", panel.first=grid())
```