### 2024-06-24 Simulation1.Rmd
Montag Abend

---
title: "Simulation"
output: pdf_document
date: "2024-06-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl) # to read the initial file
library(data.table) # for shift
```

```{r}
# Importing the database with the artificial forest
trees <- read_excel("~/work/Sim_article2024/artificial_forest_round.xls")
```

# First approach

```{r}
# First visualization
plot(x = trees$x, y = trees$y, main = "Forest")
# All trees have coordinates (abscissa and ordinate) between 200 and 800
```

```{r}
# True values
cat("Il y a au total", length(trees$x), "arbres et un volume total de", sum(trees$v), "m^3 \n", "Soit un volume moyen de", mean(trees$v), "m^3 \n", "Le diamètre moyen à hauteur de poitrine est de", mean(trees$d130), "cm")
```

# Creating cells

We may create another variable to indicate in which cell is each tree.

```{r}
# Randomizing the positions of the numbers used to identify cells
list <- sample(1:10)
```

```{r}
# Shift considered to add randomization
trend <- as.numeric(sample(1:10, 1))
```

```{r}
n <- 600 # dimension of a cell = 1 unit x 1 unit
vec <- c() # vector containing coordinates of cells
for (j in 0:9){ # for each number (1 to 10)
  for (i in 0:((n/10)-1)){ # translation
    vec <- append(vec, 200+j+10*i)
  }
}
```

```{r}
rm(i, j, n) # Useless outside loops
```

```{r}
trees$cell <- rep(0, 30942) # Initialization of the variable
```

3 steps to assign a type of cell (i.e. a number between 1 and 10) to every trees

## 1st step (1 for loop)
One line, one number, no shift yet
```{r}
for (k in 1:60) { # nombre de fois où on obtient un même identifiant pour une ligne donnée
    trees$cell <- ifelse(
    (vec[k] <= trees$x) & (trees$x < vec[k]+1) & (200 <= trees$y) & (trees$y < 200+1), 
    list[1], trees$cell)
}
```

## 2nd step (2 for loops)
All the lines, still one number
```{r}
list_stable <- list
```

```{r}
vec_test <- c()
```

```{r}
for (l in 1:599) {
  for (k in 1:60) { # nombre de fois où on obtient un même identifiant pour une ligne donnée
    vec_test <- append(vec_test, (k + (l-1)*60 -1)%%600 + 1)
    # if ((k + (l-1)*60 -1)%%600 + 1 <= 0){
    #   cat((k + (l-1)*60 -1)%%600 + 1, "k=", k, "; l=", l)
    # }
    trees$cell <- ifelse(
    (vec[(k + (l-1)*60 -1)%%600 + 1] <= trees$x) & (trees$x < vec[(k + (l-1)*60 -1)%%600 + 1]+1) & (200 + (l-1) <= trees$y) & (trees$y < 200+l), 
    # list[1 + trend*(l-1)], 
    list_stable[1],
    trees$cell)
  }
  # list <- shift(list, n = trend, type = "cyclic") # shift for each line
}
```

```{r}
# When only focusing on 1 10th of the population
plot(x = trees[trees$cell == 5,]$x, y = trees[trees$cell == 5,]$y, main = "Echantillonnage", panel.first=grid())
```

```{r}
# When only focusing on 1 10th of the population but with more precision
plot(x = trees[(trees$cell == 5) & (trees$x >= 300) & (trees$x < 310),]$x, y = trees[(trees$cell == 5) & (trees$x >= 300) & (trees$x < 310),]$y, main = "Echantillonnage", panel.first=grid())
```

## 3rd step (3 for loops)
The same thing as before but with all possible identifiers (from 1 to 10).
```{r}
for (nb in 1:10) {
  list <- list_stable
  for (l in 1:599) {
  for (k in 1:60) { # nombre de fois où on obtient un même identifiant pour une ligne donnée
    trees$cell <- ifelse(
    (vec[(k + (l-1)*60 -1)%%600 + 1 + (nb-1)*60] <= trees$x) & (trees$x < vec[(k + (l-1)*60 -1)%%600 + 1 + (nb-1)*60] + 1) & (200 + (l-1) <= trees$y) & (trees$y < 200+l), 
    list_stable[nb],
    trees$cell)
  }
  list <- shift(list, n = trend, type = "cyclic") # shift for each line
}
}
```

```{r}
# When only focusing on 1 10th of the population
plot(x = trees[trees$cell == 4,]$x, y = trees[trees$cell == 4,]$y, main = "Echantillonnage", panel.first=grid())
```

```{r}
# When only focusing on 1 10th of the population
plot(x = trees[trees$cell == 9,]$x, y = trees[trees$cell == 9,]$y, main = "Echantillonnage", panel.first=grid())
```

```{r}
# When only focusing on a small part of the population
plot(x = trees[(trees$cell == 2) & (trees$x < 350) & (trees$x >= 340),]$x, y = trees[(trees$cell == 2) & (trees$x < 350) & (trees$x >= 340),]$y, main = "Echantillonnage", panel.first=grid())
# Les points sont compris entre 341.0 et 342.0
```

```{r}
# When only focusing on a small part of the population
plot(x = trees[(trees$cell == 2) & (trees$x < 350) & (trees$x >= 340) & (trees$y >= 320) & (trees$y < 380),]$x, y = trees[(trees$cell == 2) & (trees$x < 350) & (trees$x >= 340) & (trees$y >= 320) & (trees$y < 380),]$y, main = "Echantillonnage", panel.first=grid())
# Les points sont compris entre 341.0 et 342.0
```

```{r}
# When only focusing on the null values
plot(x = trees[trees$cell == 0,]$x, y = trees[trees$cell == 0,]$y, main = "Echantillonnage", panel.first=grid())
```

```{r}
trees[trees$cell == 0,] # 4 trees that do not fill in any cell
# Common characteristic: y is very high
```

```{r}
trees[(trees$cell != 0) & (trees$y >= 799),] # vide
# Problème d'arrondi pour les y
```

# Sampling one point and "creating a circle around"

```{r}
a <- runif(1, 200, 800)
b <- runif(1, 200, 800)
```

```{r}
# Function to check if a point is inside the circle
is_point_in_circle <- function(x, y, h, k, R) {
  # Calculate the squared distance between the point and the center of the circle
  squared_distance <- (x - h)^2 + (y - k)^2
  # Check if the squared distance is less than or equal to the squared radius
  return(squared_distance <= R^2)
}
```

```{r}
trees$Is_in_circle <- is_point_in_circle(a, b, trees$x, trees$y, 5)
```

```{r}
# The sample thus obtained
trees[trees$Is_in_circle == TRUE,]
```

```{r}
# When considering only one circle
total_trees <- nrow(trees[trees$Is_in_circle == TRUE,])
total_d130 <- sum(trees[trees$Is_in_circle == TRUE,]$d130)
```

## One point at random in each cell with a given identifier

Second stage of first phase

```{r}
random_points <- data.frame(x = numeric(), y = numeric())
```

```{r}

a <- runif(1, vec[1], vec[1] + 1)
b <- runif(1, vec[1], vec[1] + 1)
```

```{r}
nb <- sample(1:10, 1)
```

```{r}
list <- list_stable
for (l in 1:599) {
  for (k in 1:60) { # nombre de fois où on obtient un même identifiant pour une ligne donnée
    a <- runif(1, vec[k + (nb-1)*60], vec[k + (nb-1)*60] + 1)
    b <- runif(1, 200 + (l-1), 200+l)
    random_points <- rbind(random_points, list(a,b))
    # trees$cell <- ifelse(
    # (vec[] <= trees$x) & (trees$x < vec[k + (nb-1)*60] + 1) & (200 + (l-1) <= trees$y) & (trees$y < 200+l), 
    # list_stable[nb],
    # trees$cell)
  }
  list <- shift(list, n = trend, type = "cyclic") # shift for each line
}

# NB: 35 940 points are randomly selected if nb = 5
# 71 880 points are randomly selected if nb = 8, idem with 10
# whereas we should have exactly 36 000 cells of each type

# PROBLEM
# N'importe quoi, ne prend pas en compte le shift
```

```{r}
# It works a priori
# l: toutes les lignes ; k: les cellules sur une ligne ; a: abscisse ; b: ordonnée

list <- list_stable
for (l in 1:600) {
  for (k in 1:60) { # nombre de fois où on obtient un même identifiant pour une ligne donnée
    colnames(random_points) <- list("x", "y")
    if (length(random_points[(random_points$x > vec[(k + (l-1)*60 -1 + (nb-1)*60)%%600 + 1])
                             & (random_points$x < vec[(k + (l-1)*60 -1 + (nb-1)*60)%%600 + 1] + 1)
                             & (random_points$y > 200 + (l-1))
                             & (random_points$y < 200 + l) ]) == 0){ # Eviter les doublons dans chaque cellule
    a <- runif(1, vec[(k + (l-1)*60 -1 + (nb-1)*60)%%600 + 1], vec[(k + (l-1)*60 -1 + (nb-1)*60)%%600 + 1] + 1)
    b <- runif(1, 200 + (l-1), 200+l)
    random_points <- rbind(random_points, list(a,b))
    }
    # trees$cell <- ifelse(
    # (vec[] <= trees$x) & (trees$x < vec[k + (nb-1)*60] + 1) & (200 + (l-1) <= trees$y) & (trees$y < 200+l), 
    # list_stable[nb],
    # trees$cell)
  }
  # list <- shift(list, n = trend, type = "cyclic") # shift for each line
}

# 36 000 points are randomly selected if nb = 5
# Which is what we want!

```

```{r}
rm(a,b,k,l) # Useless outside loops
```

```{r}
colnames(random_points) <- list("x", "y")
```

```{r}
plot(x = random_points$x, y = random_points$y, main = "Random points", panel.first=grid())
```

```{r}
# plot(x = random_points[(random_points$x < 210),]$x, y = random_points[(random_points$x < 210),]$y, main = "Random points", panel.first=grid(), xlim = c(200, 210), ylim = c(200, 220))
# Semble régulier si nb = 2
# Probleme : 2 fois trop de points tirés 
```

```{r}
# plot(x = random_points[(random_points$x < 210),]$x, y = random_points[(random_points$x < 210),]$y, main = "Random points", panel.first=grid(), xlim = c(202, 205), ylim = c(200, 205))
# Il arrive qu'il y ait 2 points dans la même cellule...
```

```{r}
plot(x = random_points[(random_points$x < 210),]$x, y = random_points[(random_points$x < 210),]$y, main = "Random points", panel.first=grid(), xlim = c(200, 210), ylim = c(200, 220))
# Semble régulier si nb = 2
```

```{r}
plot(x = random_points[(random_points$x < 210),]$x, y = random_points[(random_points$x < 210),]$y, main = "Random points", panel.first=grid(), xlim = c(202, 205), ylim = c(200, 205))
# Un seul point par cellule
```

# Circles around random points

```{r}

```

```{r}
# Adding columns to the dataframe
# Number of trees inside each circle around each random point ; Total of diameter ; Total volume
for (point in 1:36000) {
  trees$Is_in_circle <- is_point_in_circle(a, b, trees$x, trees$y, 5)
  # x, y, h, k, R
  total_trees <- nrow(trees[trees$Is_in_circle == TRUE,])
  total_d130 <- sum(trees[trees$Is_in_circle == TRUE,]$d130)
}
random_points
```

```{r}
# Creating the local variable
# Formule pages 10-11 de Chauvet et alii (2024)
M_k <- function(x, y, h, k, R) {

}
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```