### 2024-06-25 No_cell.Rmd
Martes tarde

---
title: "Uniform sampling over the all territory without any cell"
output: html_document
date: "2024-06-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("pracma") # double integral
library(readxl) # to read the initial file
library(data.table) # for shift
library(pracma)
library(magrittr)
```

# Preliminaries

```{r}
# Importing the database with the artificial forest
trees <- read_excel("~/work/Sim_article2024/artificial_forest_round.xls")
```

```{r}
# Sample size
n <- 100
```

```{r}
# Dataframe containing all coordinates of n randomly chosen points
random_points <- data.frame(x = rep(0, n), y = rep(0, n))
```

```{r}
# Uniform sampling over the whole territory
random_points$x <- runif(n, 0, 1000)
random_points$y <- runif(n, 0, 1000)
```

```{r}
# Marginal density: density of the uniform distribution over 1:1000
# Which leads us to the inclusion density function
pi <- sum(rep((1/1000)*(1/1000), n))
```

# With simple functions

## First function: (x,y) -> 1

```{r}
## First simple function: 1 given to each point
random_points$rho1 <- rep(1, n)
# Formula given in Cordy (1993), page 3
```

```{r}
## The most simple function
fun1 <- function(x,y){
  1
}
```

```{r}
# Value to estimate: the area of the territory
area <- 1000*1000
# integral2(fun1, 0, 1000, 0, 1000)$Q # the same
```

```{r}
# Estimated value
est_area <- sum((random_points$rho1)/pi) # Equal to 1 whatever the chosen points
```

## Second function: (x,y) -> (x/y) normalized

```{r}
# x/y given to each point
fun_ratio <- function(x,y){
  x/y
}
```

```{r}
# Normalization of the ratio
fun_norm <- function(x,y){
  (x/y)/(integral2(fun_ratio, 0, 1000, 0, 1000)$Q)
}
```

```{r}
random_points$rho2 <- fun_norm(random_points$x,random_points$y)
```

```{r}
# Value to estimate
ratio_norm <- integral2(fun_norm, 0, 1000, 0, 1000)$Q # = 1
```

```{r}
# Estimated value
est_ratio_norm <- sum((random_points$rho2)/pi)
# NOT SATISFACTORY: too low (btw 0 and 1)
```

```{r}
random_points$rho2 <- fun_ratio(random_points$x,random_points$y)
```

```{r}
# Value to estimate
ratio_norm <- integral2(fun_ratio, 0, 1000, 0, 1000)$Q # = 1
```

```{r}
# Estimated value
est_ratio <- sum((random_points$rho2)/pi)
# NOT SATISFACTORY
```

# Monte Carlo

```{r}
# Number of replications for the Monte-Carlo method
B <- 1000
```

```{r}
# Initialization with a first sample
series_MC <- data.frame(x.1 = runif(n, 0, 1000), y.1 = runif(n, 0, 1000))
col_names <- c("x1", "y1")
```

```{r}
# Making as many samples as necessary for MC
for (b in 2:B) {
  x <- paste("x", b, sep = "")
  y <- paste("y", b, sep = "")
  # Uniform sampling over the whole territory
  a1 <- runif(n, 0, 1000)
  a2 <- runif(n, 0, 1000)
  # series_MC$x <- runif(n, 0, 1000)
  # series_MC$y <- runif(n, 0, 1000)
  # append()
  series_MC <- series_MC %>% add_column(x = a1, y = a2)
  col_names <- append(col_names, c(x, y))
}
colnames(series_MC) <- col_names
```

```{r}
# Calculation of rho function for each pair (x,y)
rho_MC <- data.frame(x.1 = rep(0, n), y.1 = rep(0, n))
col_names <- c()
```

```{r}
for (c in 1:B) {
  col_ <- paste("est", c, sep = "")
  # With normalized ratio
  a <- fun_norm(series_MC[(2*c)-1],series_MC[2*c])
  rho_MC <- rho_MC %>% add_column(col_ = a)
  col_names <- append(col_names, col_)
}
rho_MC$x.1 <- NULL
rho_MC$y.1 <- NULL
colnames(rho_MC) <- col_names
```

```{r}
# List of estimated totals thus obtained from each sample
est_MC <- c()
```

```{r}
for (c in 1:B) {
  est_tot <- sum((rho_MC[c])/pi)
  est_MC <- append(est_MC, est_tot)
}
```

```{r}
# Empirical mean based on MC
mean(est_MC) # should be close to 1
```

```{r}
# Empirical standard deviation based on MC
sd(est_MC)
```

```{r}
### Function to automate Monte Carlo method

true_vs_MC <- function(n, B, f){
  # n: sample size; B: number of iterations for MC; f: function to consider (with 2 variables, x and y)
  pi <- sum(rep((1/1000)*(1/1000), n))

  # 1st step: generating n samples
  series_MC <- data.frame(x.1 = runif(n, 0, 1000), y.1 = runif(n, 0, 1000))
  col_names <- c("x1", "y1")
  for (b in 2:B) {
    x <- paste("x", b, sep = "")
    y <- paste("y", b, sep = "")
    # Uniform sampling over the whole territory
    a1 <- runif(n, 0, 1000)
    a2 <- runif(n, 0, 1000)
    series_MC <- series_MC %>% add_column(x = a1, y = a2)
    col_names <- append(col_names, c(x, y))
  }
  colnames(series_MC) <- col_names

  # 2nd step: creating the values corresponding to the function
  rho_MC <- data.frame(x.1 = rep(0, n), y.1 = rep(0, n))
  col_names <- c()
  for (c in 1:B) {
    col_ <- paste("est", c, sep = "")
    a <- f(series_MC[(2*c)-1],series_MC[2*c])
    rho_MC <- rho_MC %>% add_column(col_ = a)
    col_names <- append(col_names, col_)
  }
  rho_MC$x.1 <- NULL
  rho_MC$y.1 <- NULL
  colnames(rho_MC) <- col_names

  # 3rd step: List of estimated totals thus obtained from each sample
  est_MC <- c()
  for (c in 1:B) {
    est_tot <- sum((rho_MC[c])/pi)
    est_MC <- append(est_MC, est_tot)
  }

  true_value <- integral2(f, 0, 1000, 0, 1000)$Q 
  est_value <- mean(est_MC)
  return(c(true_value, est_value))
}
```

```{r}
true_vs_MC(n = 1000, B = 1000, f = fun_norm)
# 1.000000 2.265018 BOF (?)...
# 1.000000 2.127999
# 1.000000 4.984275
# 1.000000 6.072692
# 1.000000 1.982327
```

```{r}
true_vs_MC(n = 1000, B = 1000, f = fun1)
# 1e+06 1e+06 (Equality = great)
```

```{r}
true_vs_MC(n = 1000, B = 1000, f = fun_ratio)
# 3715802 6631207 BOF...
# 3715802 10413305
```

```{r}
true_vs_MC(n = 10000, B = 10000, f = fun_norm)
# 1.000000 
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```