### 2024-06-26 Test.Rmd
Mardi soir

---
title: "Test"
output:
  pdf_document: default
  html_document: default
date: "2024-06-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$\mapsto$ 1

```{r}
vec_test <- c(0,1,2)
f <- function(data, indice){
  d <- data[indice]
  return(c((d[1]+d[2]+d[3])/3))
}
boot(data=vec_test, statistic=f, R=10)
```