### 2024-07-01 Trees_no_grid.Rmd

---
title: "Uniform sampling without any cells to count trees"
output: pdf_document
date: "2024-06-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("boot") # for bootstrap
```

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column
library(boot)
```

# Preliminaries

```{r}
# Importing the database with the artificial forest
trees <- read_excel("~/work/Sim_article2024/Data/artificial_forest_round.xls")
```

For now, the variable of interest is the number of trees.

```{r}
# Sample size
n <- 100

# Radius
R <- 5

# Inclusion density function
p <- sum(rep((1/1000)*(1/1000), n))
```

# Creating one sample

A sample is created by uniformly sampling the whole territory.

```{r}
# Dataframe containing all coordinates of n randomly chosen points
# Uniform sampling over the whole territory 
random_points <- data.frame(x = runif(n, 0, 1000), y = runif(n, 0, 1000))
```

# Creating a circle around each point

We want to target trees inside the circles defined by the sample and the radius.

```{r}
# Function to check if a point is inside a circle (indicator variable)
is_point_in_circle <- function(x, y, h, k, R) {
  # Calculate the squared distance between the point and the center of the circle
  squared_distance <- (x - h)^2 + (y - k)^2
  # Check if the squared distance is less than or equal to the squared radius
  return(as.numeric(squared_distance <= R^2))
}
```

```{r}
# New dataframe called trees_in (length = number of trees)
# Each column targets one point which was previously randomly chosen
# The value is 1 if the tree corresponding is inside the circle
trees_in <- tibble(.rows = nrow(trees))
col_names <- c()
for (point in 1:n) {
  new_col <- paste("point", point, sep = "")
  a <- random_points[point,]$x
  b <- random_points[point,]$y
  val <- is_point_in_circle(a, b, trees$x, trees$y, R)
  trees_in <- trees_in %>% add_column(new_col = val)
  col_names <- append(col_names, new_col)
  colnames(trees_in) <- col_names
}
colnames(trees_in) <- col_names
```

```{r}
rm(a,b,val,col_names, point, new_col)
```

```{r}
# Creating a local variable (cf pages 10-11 in Trinh et alii (2024))
local_var <- c()
source("~/work/Sim_article2024/Inc_area.R")
for (placette in 1:n) {
  local_var <- append(local_var, sum(trees_in[trees_in[placette] != 0,]/inc_area(c(random_points[placette,]$x, random_points[placette,]$y),0,0,1000,1000,R)))
}

est_nb <- sum((local_var)/p)
cat("An estimation of the total number of trees based on the sample:", est_nb)
# 26228.73 e.g.
# 38197.19 e.g.
```

```{r}
rm(placette)
```

## Automation

A local variable is created, based on one pair of points (center of the circle) and the radius.

```{r}
circle <- function(x1, x2, R){
  # Produces a local variable based on 2 coordinates (x1, x2) and a radius (R)

  # Intermediary step: which trees are inside the circle?
  trees_in <- data.frame(useless = rep(0,30942))
  new_col <- "in_or_out"
  val <- is_point_in_circle(x1, x2, trees$x, trees$y, R)
  trees_in <- trees_in %>% add_column(new_col = val)
  trees_in[1] <- NULL

  # Creation of the local variable
  local_var <- sum(trees_in[new_col != 0,]/inc_area(c(x1, x2),0,0,1000,1000,R))
  return(local_var) # A single number
}
```

A mean of B estimations (Monte Carlo), each of them based on a sample of size n and using a radius R.

```{r}
### Function to automate Monte Carlo method

true_vs_MC <- function(n, B, R){
  # n: sample size; B: number of iterations for MC;
  # R: radius of the circle, taken into account for the sampling
  p <- sum(rep((1/1000)*(1/1000), n))

  # 1st step: generating n samples
  series_MC <- data.frame(x.1 = runif(n, 0, 1000), y.1 = runif(n, 0, 1000))
  col_names <- c("x1", "y1")
  for (b in 2:B) {
    x <- paste("x", b, sep = "")
    y <- paste("y", b, sep = "")
    # Uniform sampling over the whole territory
    a1 <- runif(n, 0, 1000)
    a2 <- runif(n, 0, 1000)
    series_MC <- series_MC %>% add_column(x = a1, y = a2)
    col_names <- append(col_names, c(x, y))
  }
  colnames(series_MC) <- col_names

  # 2nd step: creating the values corresponding to the function
  # Based on each sampled pair in all samples
  rho_MC <- tibble(.rows = n)
    # data.frame(x.1 = rep(0, n), y.1 = rep(0, n))
  col_names <- c()
  for (c in 1:B) {
    col_ <- paste("est", c, sep = "")
    a <- c()
    for (i in 1:n) {
      a <- append(a, circle(series_MC[i, (2*c)-1], series_MC[i, 2*c], R))
    }
    rho_MC <- rho_MC %>% add_column(col_ = a)
    col_names <- append(col_names, col_)
  }
  # rho_MC$x.1 <- NULL
  # rho_MC$y.1 <- NULL
  colnames(rho_MC) <- col_names

  # 3rd step: List of estimated totals thus obtained from each sample
  est_MC <- c()
  for (c in 1:B) {
    est_tot <- sum((rho_MC[c])/p)
    est_MC <- append(est_MC, est_tot)
  }

  true_value <- nrow(trees) # Theoretical value
  est_value <- mean(est_MC) # Estimated value
  return(c(true_value, est_value))
}
```

```{r}
res <- true_vs_MC(n = 100, B = 1000, R = 5)
cat("True number of trees:", res[1], "\nMean of all the estimations (Monte Carlo):", res[2])
# True number of trees: 30942 
# Mean of all the estimations (Monte Carlo): 30954.24
```