### 2024-07-09 MC_boot.Rmd
Try to merge MC and bootstrap

---
title: "Monte-Carlo and bootstrap together"
output: html_document
date: "2024-07-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column
```

# Preliminaries

```{r}
# Importing the database with the artificial forest
trees <- read_excel("~/work/Sim_article2024/Data/artificial_forest_round.xls")
```

```{r}
# Functions to produce a local variable
source("~/work/Sim_article2024/Useful_functions/Circle.R")
```

# Useful functions

```{r}
f_boot <- function(df_sample, B){
  # df_sample: original sample; B: number of bootstrap samples made thanks to df_sample

  df_boot <- tibble(.rows = nrow(df_sample))
  for (b in 1:B) {
    nb <- sample(1:nrow(df_sample), nrow(df_sample), replace = TRUE)
    sample_b <- df_sample[nb,]
    df_boot <- df_boot %>% add_column(new_col = sample_b)
    colnames(df_boot)[b] <- paste("ech", b, sep = "")
  }
  return(do.call(data.frame, df_boot))
  # Return a data frame with B samples of 2 variables and of size n
}
```

```{r}
# Making of one estimation for each sample made by the bootstrap method
est_boot <- function(df_boot, p, fun){
  # df_boot: dataframe with 2*B columns and n rows
  # fun: function used to calculate the local variable 
  # based on one point of the sample

  nboot <- length(df_boot)/2
  vec_est <- c()
  for (c in 1:nboot) {
    a <- mapply(fun, df_boot[, (2*c)-1], df_boot[, 2*c])
    vec_est <- append(vec_est, sum(a/p))
  }
  vec_est <- sort(vec_est)
  return(vec_est)
  # Vector with one estimation for each sample
}
```

# A function to plug bootstrap in a MC process

```{r}
### Function to automate Monte Carlo method

MC_boot <- function(n, B1, B2, f){
  # n: sample size; B1: number of iterations for MC;
  # B2: number of iterations for bootstrap;
  # f: function to consider (2 arguments)
  # Argument grid? (change n then)
  p <- sum(rep((1/1000)*(1/1000), n))

  est_MC <- c()
  mean_est_boot <- c()
  var_est_boot <- c()

  for (b in 2:B1) {
    # One MC iteration at a time
    random_points <- data.frame(x = runif(n, 0, 1000), y = runif(n, 0, 1000))
    a <- mapply(f, random_points[, 1], random_points[, 2])
    a <- sum(a/p)
    est_MC <- append(est_MC, a)

    # Bootstrap iterations
    df_boot <- f_boot(random_points, B2)
    est_boot_val <- est_boot(df_boot, p, f)
    mean_est_boot <- append(mean_est_boot, mean(est_boot_val))
    var_est_boot <- append(var_est_boot, var(est_boot_val))
  }
  return(c(mean(est_MC), var(est_MC), mean(var_est_boot)))
  # Estimated value of the total, Monte-Carlo variance, mean of the bootstrap variance
}
```

```{r}
MC_boot(n = 100, B1 = 1000, B2 = 200, f = circle3_nb)
```

```{r}

```