### 2024-07-11 adapt_boot2.Rmd
Not done yet

---
title: "Adapting bootstrap to our problem - Part 2"
output: pdf_document
date: "2024-07-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column
library(ggplot2)
```

The goal of this document is to create another bootstrap function that is adapted to the problem of continuous sampling with a grid. Indeed, we want a bootstrap function that takes into account the regularity of the sampling which is induced by the grid.

# Preliminaries

```{r}
# Importing the database with the artificial forest
trees <- read_excel("~/work/Sim_article2024/Data/artificial_forest_round.xls")
```

```{r}
# Functions to produce a local variable
source("~/work/Sim_article2024/Useful_functions/Circle.R")
```

# A first sample

## Sampling in one 10th of the small squares of a grid

```{r}
### Randomizing the positions of the numbers (from 1 to 10) 
# used to identify small squares in the grid
random10 <- sample(1:10)
i <- sample(1:10, 1) # choice of the rank of the identifier we want
nb <- random10[i] # the identifier thus chosen
```

```{r}
# Dimension of a small square (length of one side of the square)
# We assume that c_dim is a divisor of 1000
c_dim <- 10
```

```{r}
# Now, we sample at random following the grid
random_points <- data.frame(x = numeric(), y = numeric(), l = numeric())

# column l in addition
for (l in 0:((1000/c_dim)-1)) {
  if (l==0){
    x_init <- i*c_dim
  }else if ((3*l+i)*c_dim < (10)*c_dim){
    x_init <- (3*l+i)*c_dim
  }else{
    x_init <- (((3*l+i)*c_dim)%%(c_dim*10))
  }
  k <- 0
  while (x_init + 10*c_dim*k < 1000) {
    a <- runif(1, x_init + 10*c_dim*k, x_init + 10*c_dim*k + c_dim)
    b <- runif(1, l*c_dim, (l+1)*c_dim)
    random_points <- rbind(random_points, list(a,b, l))
    k <- k+1
  }
}
colnames(random_points) <- list("x", "y", "l")
```

```{r}
plot(x = random_points$x, y = random_points$y, xlab = "", ylab = "", main = "Random points", panel.first=grid(), las = 1)
```
Figure 1. **Points sampled following a grid over the whole territory**.

```{r}
rm(a,b,l,x_init,k)
```

```{r}
### Values available after sampling

# Sample size
n <- nrow(random_points)

# Marginal density: density of the uniform distribution over 1:1000
# Which leads us to the inclusion density function
p <- sum(rep((1/1000)*(1/1000), n))
```

## Estimations based on the sample previously made

```{r}
# Number of trees
random_points$var <- mapply(circle3_nb, random_points[, 1], random_points[, 2], MoreArgs = list(15, 9, 6))
est_init_nb <- sum(random_points$var/p)
cat("Number of trees:", nrow(trees), "\nEstimation of the number of trees:", est_init_nb)

# Ex:
# Number of trees: 30942 
# Estimation of the number of trees: 31066.18

# Number of trees: 30942 
# Estimation of the number of trees: 30489.25
```

# Bootstrap

Here, two methods are used: a naive one and an adapted one.

## First method to make bootstrap samples

```{r}
# We only keep x and y in the dataframe
random_points_1 <- random_points
random_points_1$l <- NULL
```

```{r}
f_boot <- function(df_sample, B){
  # df_sample: original sample; B: number of bootstrap samples made thanks to df_sample

  df_boot <- tibble(.rows = nrow(df_sample))
  for (b in 1:B) {
    nb <- sample(1:nrow(df_sample), nrow(df_sample), replace = TRUE)
    sample_b <- df_sample[nb,]
    df_boot <- df_boot %>% add_column(new_col = sample_b)
    colnames(df_boot)[b] <- paste("ech", b, sep = "")
  }
  return(do.call(data.frame, df_boot))
  # Return a data frame with B samples of 2 variables and of size n
}
```

## Second method to make bootstrap samples

```{r}
f_boot2 <- function(df_sample, nboot){
  # df_sample: original sample (4 variables: x,y,l,var)
  # nboot: number of bootstrap samples made thanks to df_sample

  odd_even1 <- sample(1:2, 1) # Randomly put more weight on even or odd lines
  odd_even2 <- sample(1:2, 1) # The same for "columns"

  df_sample$step <- rep(0, nrow(df_sample))
  for (line in unique(df_sample$l)) {
    df_sample$step <- as.numeric(df_sample$step == 0)*(row(df_sample[df_sample$l == line,])[,1]%%2) + as.numeric(df_sample$step != 0)*df_sample$step
  }

  df_sample$prob <- (((df_sample$l+odd_even1)%%2 + (df_sample$step+odd_even2)%%2) == 0)*(1/4) +
    (((df_sample$l+odd_even1)%%2 + (df_sample$step+odd_even2)%%2) == 1)*(1/2) +
    (((df_sample$l+odd_even1)%%2 + (df_sample$step+odd_even2)%%2) == 2)*(3/4)
  df_sample$l <- NULL
  df_sample$step <- NULL

  df_boot <- tibble(.rows = nrow(df_sample))
  for (b in 1:nboot) {
    nb <- sample(1:nrow(df_sample), nrow(df_sample), replace = TRUE, prob = df_sample$prob)
    df_sample$prob <- NULL
    sample_b <- df_sample[nb,]
    df_boot <- df_boot %>% add_column(new_col = sample_b)
    colnames(df_boot)[b] <- paste("ech", b, sep = "")
  }
  return(do.call(data.frame, df_boot))
  # Return a data frame with nboot samples of 2 variables and of size n
}
```

## Number of trees

```{r}
# Making of one estimation for each sample made by the bootstrap method
est_boot1 <- function(df){
  # df: dataframe with 3*B columns and n rows

  nboot <- length(df)/3
  vec_est <- c()
  for (c in 1:nboot) {
    vec_est <- append(vec_est, sum(df[,3*c]/p))
  }
  vec_est <- sort(vec_est)
  return(vec_est)
  # Vector with one estimation for each sample
}
```

# Comparison between both methods

```{r}
comparison2 <- function(B) {
  # B: number of bootstrap replications each time
  mean_vec1 <- c()
  mean_vec2 <- c()
  var_vec1 <- c()
  var_vec2 <- c()
  b1 <- 100
  while (b1  < B) {
    ## First method
    # We only keep x and y in the dataframe
    random_points_1 <- random_points
    random_points_1$l <- NULL
    df_boot <- f_boot(random_points_1, b1)
    est_boot_1 <- est_boot1(df_boot)

    ## Second method
    df_boot2 <- f_boot2(random_points, b1)
    est_boot_2 <- est_boot1(df_boot2)

    mean_vec1 <- append(mean_vec1, abs(mean(est_boot_1)-30942))
    mean_vec2 <- append(mean_vec2, abs(mean(est_boot_2)-30942))

    var_vec1 <- append(var_vec1, var(est_boot_1))
    var_vec2 <- append(var_vec2, var(est_boot_2))

    # tot_mean <- append(tot_mean, as.numeric(abs(mean(est_boot_2)-30942) <  abs(mean(est_boot_1)-30942))*2-1)
    # 1: the second method is better
    # tot_var <- append(tot_var, as.numeric(var(est_boot_2) <  var(est_boot_1))*2-1)
    # 1: the second method is better
    b1 <- b1+10
  }
  return(list(mean_vec1, mean_vec2, var_vec1, var_vec2))
}
```

```{r}
res2 <- comparison2(500)
```

```{r}
res <- comparison2(500)
```

```{r}
data = as.data.frame(cbind(1:length(res2[[1]]), res2[[1]], res2[[2]]))
p1 <- ggplot(data, aes(data[,1], data[,2]))
p2 <- ggplot(data, aes(data[,1], data[,3]))
p1 + p2 + geom_point()
p1 + geom_point

# p <- ggplot(data, aes(data[,1], cbind(data[,2], data[,3])))
# p + geom_point()

p <- ggplot(data, aes(data[,1], data[,3] - data[,2]))
p + geom_point()
```

```{r}
data_var = as.data.frame(cbind(1:length(res2[[1]]), res2[[3]], res2[[4]]))

p <- ggplot(data_var, aes(data_var[,1], data_var[,3] - data_var[,2]))
p + geom_point()
```

```{r}
data_var = as.data.frame(cbind(1:length(res2[[1]]), res2[[3]], res2[[4]]))

p <- ggplot(data_var, aes(data_var[,1], data_var[,3] - data_var[,2]))
p + geom_point()
```

```{r}
ggplot(x = 1:length(res[[1]]), y = res[[1]]) +
  plot(x = 1:length(res[[1]]), y = res[[2]], col = "red")
geom_point(
    mapping = aes(y = group_mean), data = group_means_df,
    colour = 'red', size = 3)

```

```{r}
# If it is positive, the second method is the best to minimize the variance
sum(res[[2]])
# With ncomp = 100 and B = 500
# 2 e.g.
# 12

# With ncomp = 1000 and B = 800
# 56
```

# Conclusion