### 2024-07-15 Two_var.Rmd
Two_var cleaner

---
title: "Making sure the variances converge"
output:
  pdf_document: default
  html_document: default
date: "2024-07-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column
```

# Preliminaries

```{r}
# Importing the database with the artificial forest
trees <- read_excel("~/work/Sim_article2024/Data/artificial_forest_round.xls")
```

```{r}
# Functions to produce a local variable
source("~/work/Sim_article2024/Useful_functions/Circle.R")
```

```{r}
# Sample size
n <- 1000
```

# Monte Carlo alone

10 000 Monte Carlo iterations would be fine

```{r}
### Function to automate Monte Carlo method

MonteCarlo <- function(n, B, f){
  # n: sample size; B: number of iterations for MC; 
  # f: function to consider (with 2 variables, x and y)
  p <- sum(rep((1/1000)*(1/1000), n))

  # 1st step: generating n samples
  # series_MC <- data.frame(x.1 = runif(n, 0, 1000), y.1 = runif(n, 0, 1000))
  col_names <- c()

  series_MC <- data.frame(rep(0,n))
  est_MC <- c()
  for (b in 2:B) {
    x <- paste("x", b, sep = "")
    y <- paste("y", b, sep = "")
    z <- paste("est", b, sep = "")
    # Uniform sampling over the whole territory
    a1 <- runif(n, 0, 1000)
    a2 <- runif(n, 0, 1000)
    a <- mapply(f, a1, a2)
    est_MC <- append(est_MC, sum(a/p))
    series_MC <- series_MC %>% add_column(x = a1, y = a2, z = a)
    col_names <- append(col_names, c(x, y, z))
  }
  series_MC <- series_MC[,-1]
  colnames(series_MC) <- col_names

  return(c(mean(est_MC), var(est_MC)))
  # Estimated value and Monte-Carlo variance
}
```

```{r}
start.time <- Sys.time()

MC1 <- MonteCarlo(n, 10000, circle_nb)
# Just one circle to begin with

end.time <- Sys.time()
time.taken <- end.time - start.time

time.taken
```

```{r}
cat("Monte Carlo variance:", MC1[2])
```

# Bootstrap for each Monte Carlo iteration

1 000 Monte Carlo iterations should be fine
400 bootstrap iterations for each would be fine (200 to begin with if it is too much)

```{r}
f_boot <- function(df_sample, B){
  # df_sample: original sample; B: number of bootstrap samples made thanks to df_sample

  df_boot <- tibble(.rows = (nrow(df_sample)-1))
  for (b in 1:B) {
    nb <- sample(1:(nrow(df_sample)-1), nrow(df_sample), replace = TRUE)
    # (n-1) points at random
    sample_b <- df_sample[nb,]
    df_boot <- df_boot %>% add_column(new_col = sample_b)
    colnames(df_boot)[b] <- paste("ech", b, sep = "")
  }
  return(do.call(data.frame, df_boot))
  # Return a data frame with B samples of 2 variables and of size n
}
```

```{r}
# Making of one estimation for each sample made by the bootstrap method
est_boot <- function(df, p){
  # df: dataframe with 3*B columns and n rows

  nboot <- length(df)/3
  vec_est <- c()
  for (c in 1:nboot) {
    vec_est <- append(vec_est, sum(df[,3*c]/p))
  }
  vec_est <- sort(vec_est)
  return(vec_est)
  # Vector with one estimation for each sample
}
```

```{r}
### Function to automate Monte Carlo method

MC_boot <- function(n, B1, B2, f){
  # n: sample size; B1: number of iterations for MC;
  # B2: number of iterations for bootstrap;
  # f: function to consider (2 arguments)
  # Argument grid? (change n then)
  p <- sum(rep((1/1000)*(1/1000), n))

  est_MC <- c()
  mean_est_boot <- c()
  var_est_boot <- c()

  for (b in 2:B1) {
    # One MC iteration at a time
    random_points <- data.frame(x = runif(n, 0, 1000), y = runif(n, 0, 1000))
    random_points$var <- mapply(f, random_points[, 1], random_points[, 2])
    a <- sum(random_points$var/p)
    est_MC <- append(est_MC, a)

    # Bootstrap iterations
    df_boot <- f_boot(random_points, B2)
    est_boot_val <- est_boot(df_boot, p)
    mean_est_boot <- append(mean_est_boot, mean(est_boot_val))
    var_est_boot <- append(var_est_boot, var(est_boot_val))
  }
  return(c(mean(est_MC), var(est_MC), mean(var_est_boot)))
  # Estimated value of the total, Monte-Carlo variance, mean of the bootstrap variance
}
```

```{r}
start.time <- Sys.time()

res <- MC_boot(n, B1 = 1000, B2 = 400, f = circle_nb)
# Just one circle to begin with

end.time <- Sys.time()
time.taken <- end.time - start.time

time.taken
```

```{r}
cat("Mean of the variance of bootstrap samples:", res[3])
```