### 2024-07-23 Ratios.Rmd
Ratio with example

---
title: "Various ratios and their estimations"
output: pdf_document
date: "2024-07-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column
```

# Preliminaries

```{r}
# Importing the database with the artificial forest
trees <- read_excel("~/work/Sim_article2024/Data/artificial_forest_round.xls")
```

```{r}
# Functions to produce a local variable
source("~/work/Sim_article2024/Useful_functions/Circle.R")
```

```{r}
# Sample size
n <- 1000

# Inclusion density
p <- sum(rep((1/1000)*(1/1000), n))
```

```{r}
random_points <- data.frame(x = runif(n, 0, 1000), y = runif(n, 0, 1000))
```

# Bootstrap for each Monte Carlo iteration

1 000 Monte Carlo iterations and 400 bootstrap iterations for each one are appropriated.

```{r}
f_boot <- function(df_sample, B){
  # df_sample: original sample; 
  # B: number of bootstrap samples made thanks to df_sample

  df_boot <- tibble(.rows = (nrow(df_sample)-1))
  for (b in 1:B) {
    nb <- sample(1:nrow(df_sample), (nrow(df_sample)-1), replace = TRUE)
    # (n-1) points at random
    sample_b <- df_sample[nb,]
    df_boot <- df_boot %>% add_column(new_col = sample_b)
    colnames(df_boot)[b] <- paste("ech", b, sep = "")
  }
  return(do.call(data.frame, df_boot))
  # Return a data frame with B samples of 2 (or more) variables and of size n
}
```

```{r}
# Making of one estimation for each sample made by the bootstrap method
est_boot <- function(df, p){
  # df: dataframe with 4*B columns and n rows

  n <- nrow(df) + 1 # sample size of the initial sample
  nboot <- length(df)/4 # number of bootstrap samples
  vec_est1 <- c()
  vec_est2 <- c()
  for (c in 1:nboot) {
    vec_est1 <- append(vec_est1, (n/(n-1))*sum(df[,3*c]/p))
    vec_est2 <- append(vec_est2, (n/(n-1))*sum(df[,4*c]/p))
    # (n/(n-1)) is the correction needed since we choose only (n-1) points
  }
  # vec_est1 <- sort(vec_est1)
  # vec_est2 <- sort(vec_est2)
  vec_est <- data.frame(est1 = vec_est1, est2 = vec_est2)
  return(vec_est)
  # Vector with one estimation for each sample
}
```

```{r}
### Function to automate Monte Carlo + bootstrap method

MC_boot <- function(n, B1, B2, f1, f2){
  # n: sample size; B1: number of iterations for MC;
  # B2: number of iterations for bootstrap;
  # f1: function to consider (2 arguments) on the numerator
  # f2: function to consider (2 arguments) on the denominator

  p <- sum(rep((1/1000)*(1/1000), n))
  var_est_boot <- c()
  for (b in 2:B1) {
    # One MC iteration at a time
    random_points <- data.frame(x = runif(n, 0, 1000), y = runif(n, 0, 1000))
    random_points$var1 <- mapply(f1, random_points[, 1], random_points[, 2])
    random_points$var2 <- mapply(f2, random_points[, 1], random_points[, 2])

    # Bootstrap iterations
    df_boot <- f_boot(random_points, B2)
    est_boot_val <- est_boot(df_boot, p)
    est_boot_ratio <- est_boot_val[,1]/est_boot_val[,2]
    var_est_boot <- append(var_est_boot, var(est_boot_ratio))
  }
  return(mean(var_est_boot))
  # Mean of the bootstrap variance
}
```

# Global estimations with one circle

```{r}
tot_nb <- mapply(circle_nb, random_points[, 1], random_points[, 2])
tot_vol <- mapply(circle_vol, random_points[, 1], random_points[, 2])
```

```{r}
tot_nb_vol <- data.frame(nb = tot_nb, vol = tot_vol)
a1 <- sum(tot_nb/p)
a2 <- sum(tot_vol/p)
est_ratio <- a2/a1
```

```{r}
cat("True ratio \"total volume over number of trees\":", sum(trees$v)/nrow(trees), "\nEstimation of this ratio:", est_ratio)
# Ex:
# True ratio "total volume over number of trees": 0.1749312 
# Estimation of this ratio: 0.1714633
```

```{r}
start.time <- Sys.time()

res_ratio1 <- MC_boot(n, B1 = 1000, B2 = 400, f1 = circle_vol, f2 = circle_nb)
# Just one circle to begin with

end.time <- Sys.time()
time.taken <- end.time - start.time

time.taken
```

```{r}
cat("Mean of the variance of bootstrap samples:", res_ratio1)
# Ex:
# Mean of the variance of bootstrap samples: 2190764978
```

# Various circumferences

```{r}
y1 <- mapply(circle_vol_size, random_points[, 1], random_points[, 2])
Y1 <- sum(y1/p)
```

```{r}
cat("True ratio \"volume of the smallest over total volume\":", sum(trees[(trees$d130 > 23.5/pi) & (trees$d130 < 70.5/pi),]$v)/sum(trees$v), "\nEstimation of this ratio:", Y1/a2)
# Ex:
# True ratio "volume of the smallest over total volume": 0.5741108
# Estimation of this ratio: 0.5846677
```

```{r}
start.time <- Sys.time()

res_ratio2 <- MC_boot(n, B1 = 1000, B2 = 400, f1 = circle_vol_size, f2 = circle_vol)
# Just one circle to begin with

end.time <- Sys.time()
time.taken <- end.time - start.time

time.taken
```

```{r}
cat("Mean of the variance of bootstrap samples:", res_ratio2)
```

```{r}
y2 <- mapply(circle_vol_size, random_points[, 1], random_points[, 2], MoreArgs = list(5, c(70.5, 117.5)))
Y2 <- sum(y2/p)
```

```{r}
cat("True ratio \"volume of the medium over total volume\":", sum(trees[(trees$d130 > 70.5/pi) & (trees$d130 < 117.5/pi),]$v)/sum(trees$v), "\nEstimation of this ratio:", Y2/a2)
# Ex:
# True ratio "volume of the medium over total volume": 0.3628267
# Estimation of this ratio: 0.3623958
```

```{r}
y3 <- mapply(circle_vol_size, random_points[, 1], random_points[, 2], MoreArgs = list(5, c(117.5, 500)))
Y3 <- sum(y3/p)
```

```{r}
cat("True ratio \"volume of the biggest over total volume\":", sum(trees[(trees$d130 > 117.5/pi) & (trees$d130 < 500/pi),]$v)/sum(trees$v), "\nEstimation of this ratio:", Y3/a2)
# True ratio "volume of the biggest over total volume": 0.06306254
# Estimation of this ratio: 0.0529365
```