### 2024-07-25 Test_dens.Rmd
More global MC and function for local variables

---
title: "Test_dens"
output: pdf_document
date: "2024-07-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of this file is to correct a mistake made on previous versions of functions to create local variables (those previous versions may be found in Old_Circle.R).

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column and tibble
```

# Preliminaries

```{r}
# Importing the database with the artificial forest
trees <- read_excel("~/work/Sim_article2024/Data/artificial_forest_round.xls")
```

The variable of interest is the number of trees.

```{r}
# Sample size
n <- 100

# Radius
R <- 4
```

# Two functions

```{r}
## Function to produce the local variable for the number of trees
circle_nb_old <- function(x1, x2, R=4){
  # 2 coordinates (x1, x2) and a radius (R)

  # Preliminary: inclusion area
  source("~/work/Sim_article2024/Useful_functions/Inc_area.R")

  list_trees <- which(((x1 - trees$x)^2 + (x2 - trees$y)^2 <= R^2) & (trees$x > frontier[1]) & (trees$x < frontier[3]) & (trees$y > frontier[2]) & (trees$y < frontier[4]))
  local_var <- sum(length(list_trees)/inc_area(c(x1, x2),frontier[1],frontier[2],frontier[3],frontier[4],R))
  return(local_var)
  # list_trees: identifiers of targeted trees; local_var: local variable based on the number of trees
}
```

```{r}
## Function to produce the local variable for the number of trees
circle_nb_new2 <- function(x1, x2, R=4){
  # 2 coordinates (x1, x2) and a radius (R)

  # Preliminary: inclusion area
  source("~/work/Sim_article2024/Useful_functions/Inc_area_2.R")

  list_trees <- which(((x1 - trees$x)^2 + (x2 - trees$y)^2 <= R^2) & (trees$x > frontier[1]) & (trees$x < frontier[3]) & (trees$y > frontier[2]) & (trees$y < frontier[4]))
  trees1 <- trees
  trees1$var1 <- rep(1,nrow(trees1))
  if (length(list_trees) != 0){
    local_var <- sum((trees1[list_trees,]$var1)/mapply(inc_area_test, trees1[list_trees,]$x, trees1[list_trees,]$y, MoreArgs = list(frontier[1],frontier[2],frontier[3],frontier[4],R)))
  }
  else{local_var <- 0}
  return(local_var)
  # list_trees: identifiers of targeted trees; local_var: local variable based on the number of trees
}
```

# Tests

```{r}
# Small square
frontier <- c(500, 568, 510, 578)
# Ex: frontier <- c(500, 516, 510, 526)
# frontier <- c(500, 557, 510, 567)
# frontier <- c(500, 568, 510, 578)
```

```{r}
plot(trees$x, trees$y, xlim = c(frontier[1], frontier[3]), ylim = c(frontier[2], frontier[4]))
```

```{r}
true_val <- nrow(trees[(trees$x > frontier[1]) & (trees$x < frontier[3]) & (trees$y > frontier[2]) & (trees$y < frontier[4]),])
true_val
# Ex: 19
# 7
# 12
```

```{r}
# Inclusion density function
p1 <- (1/(frontier[3] - frontier[1]))*(1/(frontier[4] - frontier[2]))
p <- sum(rep(p1, n))
```

```{r}
gap_old <- c()
gap_new <- c()
i = 1
for (i in 1:1000) {
  sampling <- data.frame(x.1 = runif(n, frontier[1], frontier[3]), y.1 = runif(n, frontier[2], frontier[4]))
  est_old <- sum(mapply(circle_nb_old, sampling[, 1], sampling[, 2])/p)
  est_new <- sum(mapply(circle_nb_new2, sampling[, 1], sampling[, 2])/p)
  gap_old <- append(gap_old, abs(est_old-true_val))
  gap_new <- append(gap_new, abs(est_new-true_val))
  i <- i+1
}
```

```{r}
sum(gap_new - gap_old)
# Ex: -0.6188775
# -546.3402
# -97.20414
```

```{r}
sum(as.numeric(gap_new > gap_old) - as.numeric(gap_new < gap_old))
# Ex: 8
# -918
# -188
```