### 2024-07-25 MC_all.Rmd
Obviously a pb with variances

---
title: "Monte Carlo for all parameters"
output: pdf_document
date: "2024-07-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column
```

# Preliminaries

```{r}
# Importing the database with the artificial forest
all_trees <- read_excel("~/work/Sim_article2024/Data/artificial_forest_round.xls")
# To get closer to the NFI process, trees whose circumference is under 23.5 cm are not considered
trees <- all_trees[(pi*all_trees$d130 > 23.5),]
```

```{r}
# Functions to produce a local variable
source("~/work/Sim_article2024/Useful_functions/Circle.R")
```

```{r}
# Sample size
n <- 1000
```

```{r}
# True values
cat("Number of trees:", length(trees$x), "\nTotal volume:", sum(trees$v), "\nWhich leads to the following ratio \"Total volume / Number of trees\":", sum(trees$v)/nrow(trees))
```

# Monte Carlo alone

10 000 Monte Carlo iterations are appropriated.

```{r}
### Function to automate Monte Carlo method, to produce a variance

MonteCarlo <- function(n, B){
  # n: sample size;
  # B: number of iterations for MC

  p <- sum(rep((1/1000)*(1/1000), n)) # inclusion density

  MC_df <- data.frame(useless = rep(0, n))
  MC_est <- data.frame(useless = rep(0, 3), row.names = c("est_nb", "est_vol", "est_ratio"))

  for (b in 1:B) { # All in one "for" loop
    # Uniform sampling over the whole territory
    a1 <- runif(n, 0, 1000)
    a2 <- runif(n, 0, 1000)
    # Variables of interest derived from the sample
    var <- mapply(circle_all, a1, a2)
    MC_b <- data.frame(x = a1, y = a2, z = t(var)[,1], t = t(var)[,2])
    MC_df <- cbind(MC_df, MC_b)
    colnames(MC_df)[(4*(b-1)+2):(4*(b-1)+2+3)] <- c(paste("x", b, sep = ""), paste("y", b, sep = ""), paste("nb", b, sep = ""), paste("vol", b, sep = ""))
    est_b <- t(data.frame(x = sum(MC_df[4*(b-1)+2+2] /p), y = sum(MC_df[4*(b-1)+2+3] /p), z = (sum(MC_df[4*(b-1)+2+3] /p))/(sum(MC_df[4*(b-1)+2+2]/p))))
    MC_est <- cbind(MC_est, est_b)
    colnames(MC_est)[b+1] <- paste("ech", b, sep = "") 
  }
  MC_df$useless <- NULL
  MC_est$useless <- NULL
  MC_var <- c(var(as.vector(t(MC_est[1,]))), var(as.vector(t(MC_est[2,]))), var(as.vector(t(MC_est[3,]))))

  # Confidence intervals
  norm_CI <- data.frame(useless = rep(0, 2))
  perc_CI <- data.frame(useless = rep(0, 2))
  rev_perc_CI <- data.frame(useless = rep(0, 2))
  for (l in 1:3) { # Loop: one estimator each time
    est_init <- as.vector(t(MC_est[l,]))[1]
    sorted_est <- sort(as.vector(t(MC_est[l,])))
    lim_inf <- sorted_est[floor(B*2.5/100)]
    lim_sup <- sorted_est[B - floor(B*2.5/100)]
    perc_CI <- cbind(perc_CI, data.frame(c(lim_inf, lim_sup)))
    rev_perc_CI <- cbind(rev_perc_CI, data.frame(c(est_init*2 - lim_sup, est_init*2 - lim_inf)))
    norm_CI <- cbind(norm_CI, data.frame(c(est_init-qnorm(0.975)*sqrt(MC_var[l]), est_init+qnorm(0.975)*sqrt(MC_var[l]))))
  }
  norm_CI$useless <- NULL
  perc_CI$useless <- NULL
  rev_perc_CI$useless <- NULL
  colnames(norm_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
  colnames(perc_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
  colnames(rev_perc_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
  all_CI <- list(Normal = norm_CI, Percentile = perc_CI, `Reverse percentile` = rev_perc_CI)

  return(list(`Local Variables` = MC_df, Estimations = MC_est, Variances = MC_var, `Confidence Intervals` = all_CI))
  # Local Variables: local variables (number and volume) for each point in each sample
  # Estimations: 3 Estimations (number, volume, ratio volume / number) for each sample
  # Variances: Monte Carlo variance for each of the totals / ratio
  # Confidence Intervals: 3 different forms of 95% confidence intervals, for each total / ratio
}
```

# Bootstrap for each Monte Carlo iteration

1 000 Monte Carlo iterations and 400 bootstrap iterations for each one are appropriated.

```{r}
f_boot <- function(df_sample, B){
  # df_sample: original sample; 
  # B: number of bootstrap samples made thanks to df_sample

  df_boot <- tibble(.rows = (nrow(df_sample)-1))
  for (b in 1:B) {
    nb <- sample(1:nrow(df_sample), (nrow(df_sample)-1), replace = TRUE)
    # (n-1) points at random
    sample_b <- df_sample[nb,]
    df_boot <- df_boot %>% add_column(new_col = sample_b)
    colnames(df_boot)[b] <- paste("ech", b, sep = "")
  }
  return(do.call(data.frame, df_boot))
  # Return a data frame with B samples of as many variables as we want (in this application, 4) and of size (n-1)
}
```

```{r}
# Making of estimations for each sample made by the bootstrap method
est_boot <- function(df, p){
  # df: dataframe with 4*B columns and n rows

  n <- nrow(df) + 1 # sample size of the initial sample
  nboot <- length(df)/4 # number of bootstrap samples
  vec_est1 <- c()
  vec_est2 <- c()
  vec_est3 <- c()
  for (c in 1:nboot) {
    vec_est1 <- append(vec_est1, (n/(n-1))*sum(df[,3*c]/p))
    vec_est2 <- append(vec_est2, (n/(n-1))*sum(df[,4*c]/p))
    vec_est3 <- append(vec_est3, ((n/(n-1))*sum(df[,4*c]/p))/((n/(n-1))*sum(df[,3*c]/p)))
    # (n/(n-1)) is the correction needed since we choose only (n-1) points
  }
  vec_est <- data.frame(est_nb = vec_est1, est_vol = vec_est2, est_ratio = vec_est3)
  return(vec_est)
  # Vector with 3 estimations for each sample
}
```

```{r}
### Function to automate Monte Carlo + bootstrap method

MC_boot <- function(n, B1, B2){
  # n: sample size;
  # B1: number of iterations for MC;
  # B2: number of iterations for bootstrap

  p <- sum(rep((1/1000)*(1/1000), n))
  BOOT_df <- list() # All local variables
  BOOT_est <- list() # All estimations
  BOOT_var <- list() # Variances
  BOOT_CI <- list() # Confidence Intervals
  for (b in 1:B1) { # One MC iteration at a time
    # Uniform sampling over the whole territory
    a1 <- runif(n, 0, 1000)
    a2 <- runif(n, 0, 1000)
    var <- mapply(circle_all, a1, a2)
    df_b <- data.frame(x = a1, y = a2, z = t(var)[,1], t = t(var)[,2])
    colnames(df_b) <- c(paste("x.MC", b, sep = ""), paste("y.MC", b, sep = ""), paste("nb.MC", b, sep = ""), paste("vol.MC", b, sep = ""))
    est_b <- t(data.frame(x = sum(df_b[3] /p), y = sum(df_b[4] /p), z = (sum(df_b[4] /p))/(sum(df_b[3]/p))))

    # Bootstrap iterations
    df_boot <- f_boot(df_b, B2)
    est_boot_val <- est_boot(df_boot, p)

    BOOT_var[[b]] <- list(var_nb = var(est_boot_val$est_nb), var_vol = var(est_boot_val$est_vol), var_ratio = var(est_boot_val$est_ratio))
    names(BOOT_var)[b] <- paste("Variance", b, sep = "")
    BOOT_df <- c(BOOT_df, list(data_init = df_b, data_boot = df_boot))
    names(BOOT_df)[(2*b-1):(2*b)] <- c(paste("df_MC", b, sep = ""), paste("df_boot", b, sep = ""))
    BOOT_est <- c(BOOT_est, list(est_init_b = est_b, est_boot_b = est_boot_val))
    names(BOOT_est)[(2*b-1):(2*b)] <- c(paste("est_MC", b, sep = ""), paste("est_boot", b, sep = ""))

    # Confidence intervals
    norm_CI <- data.frame(useless = rep(0, 2))
    perc_CI <- data.frame(useless = rep(0, 2))
    rev_perc_CI <- data.frame(useless = rep(0, 2))
    for (l in 1:3) { # Loop: one estimator each time
      sorted_est <- sort(est_boot_val[,l])
      lim_inf <- sorted_est[floor(B2*2.5/100)]
      lim_sup <- sorted_est[B2 - floor(B2*2.5/100)]
      perc_CI <- cbind(perc_CI, data.frame(c(lim_inf, lim_sup)))
      rev_perc_CI <- cbind(rev_perc_CI, data.frame(c(est_b[l,]*2 - lim_sup, est_b[l,]*2 - lim_inf)))
      norm_CI <- cbind(norm_CI, data.frame(c(est_b[l,]-qnorm(0.975)*sqrt(var(est_boot_val[,l])), est_b[l,]+qnorm(0.975)*sqrt(var(est_boot_val[,l])))))
    }
    norm_CI$useless <- NULL
    perc_CI$useless <- NULL
    rev_perc_CI$useless <- NULL
    colnames(norm_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
    colnames(perc_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
    colnames(rev_perc_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
    all_CI <- list(Normal = norm_CI, Percentile = perc_CI, `Reverse percentile` = rev_perc_CI)
    # BOOT_CI <- c(BOOT_CI, CI_b = all_CI)
    BOOT_CI[[b]] <- all_CI
    names(BOOT_CI)[b] <- paste("CI", b, sep = "")
  }
  return(list(`Local Variables` = BOOT_df, Estimations = BOOT_est, Variances = BOOT_var, `Confidence Intervals` = BOOT_CI))
  # Local Variables: local variables (number and volume) for each point in each sample
  # Estimations: 3 estimations (number, volume, ratio volume / number) for each sample
  # Variances: mean of the bootstrap variance for each estimator
  # Confidence Intervals: 3 different forms of 95% confidence intervals, for each total / ratio
}
```

# Example

## Monte Carlo alone

```{r}
start.time <- Sys.time()

res1 <- MonteCarlo(n, 10000)
# Just one circle to begin with

end.time <- Sys.time()
time.taken <- end.time - start.time

time.taken
```

```{r}
cat("Monte Carlo means of estimations: \n - for the number of trees:", mean(as.vector(t(res1$Estimations[1,]))), "\n - for the volume of trees:", mean(as.vector(t(res1$Estimations[2,]))), "\n - for the ratio volume / number of trees:", mean(as.vector(t(res1$Estimations[3,]))))

# Example:
# Monte Carlo means of estimations: 
#  - for the number of trees: 29299.58 
#  - for the volume of trees: 5418.03 
#  - for the ratio volume / number of trees: 0.1849162
```

```{r}
cat("Monte Carlo variances: \n - for the number of trees:", res1$Variances[1], "\n - for the volume of trees:", res1$Variances[2], "\n - for the ratio volume / number of trees:", res1$Variances[3])

# Example:
# Monte Carlo variances: 
#  - for the number of trees: 2531530 
#  - for the volume of trees: 96840.38 
#  - for the ratio volume / number of trees: 1.116039e-05
```

```{r}
cat("95% confidence intervals using the normal method:\n - for the number of trees: [", res1$`Confidence Intervals`$Normal[1,1], ";", res1$`Confidence Intervals`$Normal[2,1], 
    "]\n - for the volume of trees: [", res1$`Confidence Intervals`$Normal[1,2], ";", res1$`Confidence Intervals`$Normal[2,2], 
    "]\n - for the ratio volume / number of trees: [", res1$`Confidence Intervals`$Normal[1,3], ";", res1$`Confidence Intervals`$Normal[2,3],
    "]")

# Example:
# 95% confidence intervals using the normal method: 
#  - for the number of trees: [ 25885.94 ; 32122.85 ]
#  - for the volume of trees: [ 4727.737 ; 5947.587 ]
#  - for the ratio volume / number of trees: [ 0.1774817 ; 0.1905771 ]
```

```{r}
cat("95% confidence intervals using the percentile method:\n - for the number of trees: [", res1$`Confidence Intervals`$Percentile[1,1], ";", res1$`Confidence Intervals`$Percentile[2,1], 
    "]\n - for the volume of trees: [", res1$`Confidence Intervals`$Percentile[1,2], ";", res1$`Confidence Intervals`$Percentile[2,2], 
    "]\n - for the ratio volume / number of trees: [", res1$`Confidence Intervals`$Percentile[1,3], ";", res1$`Confidence Intervals`$Percentile[2,3],
    "]")

# Example:
# 95% confidence intervals using the percentile method: 
#  - for the number of trees: [ 26203.27 ; 32429.41 ]
#  - for the volume of trees: [ 4817.327 ; 6027.707 ]
#  - for the ratio volume / number of trees: [ 0.1786038 ; 0.1915195 ]
```

```{r}
cat("95% confidence intervals using the reverse percentile method:\n - for the number of trees: [", res1$`Confidence Intervals`$`Reverse percentile`[1,1], ";", res1$`Confidence Intervals`$`Reverse percentile`[2,1], 
    "]\n - for the volume of trees: [", res1$`Confidence Intervals`$`Reverse percentile`[1,2], ";", res1$`Confidence Intervals`$`Reverse percentile`[2,2], 
    "]\n - for the ratio volume / number of trees: [", res1$`Confidence Intervals`$`Reverse percentile`[1,3], ";", res1$`Confidence Intervals`$`Reverse percentile`[2,3],
    "]"
    )

# Example:
# 95% confidence intervals using the reverse percentile method:
#  - for the number of trees: [ 25579.38 ; 31805.52 ]
#  - for the volume of trees: [ 4647.617 ; 5857.997 ]
#  - for the ratio volume / number of trees: [ 0.1765393 ; 0.189455 ]
```

## Monte Carlo and bootstrap

```{r}
start.time <- Sys.time()

res2 <- MC_boot(n, B1 = 1000, B2 = 400)
# Just one circle to begin with

end.time <- Sys.time()
time.taken <- end.time - start.time

time.taken
```

```{r}
VAR_nb <- c()
for (elt in res2$Variances) {
  VAR_nb <- append(VAR_nb, elt$var_nb)
}

VAR_vol <- c()
for (elt in res2$Variances) {
  VAR_vol <- append(VAR_vol, elt$var_vol)
}

VAR_ratio <- c()
for (elt in res2$Variances) {
  VAR_ratio <- append(VAR_ratio, elt$var_ratio)
}
```

```{r}
cat("Mean of the variance of bootstrap samples:\n - for the number of trees:", mean(VAR_nb), "\n - for the volume of trees:", mean(VAR_vol), "\n - for the ratio volume / number of trees:", mean(VAR_ratio))

# Example:
# Mean of the variance of bootstrap samples:
#  - for the number of trees: 6.273237e+16 
#  - for the volume of trees: 98861.68 
#  - for the ratio volume / number of trees: 0.1741024
```

## Comparison of the variances

```{r}
diff_percent <- function(x,y)100*(max(x,y)-min(x,y))/min(x,y)
cat("Difference between the two estimations of variance: \n - for the number of trees: ", diff_percent(res1$Variances[1], mean(VAR_nb)), "%", "\n - for the volume of trees: ", diff_percent(res1$Variances[2], mean(VAR_vol)), "%\n - for the ratio volume / number of trees: ", diff_percent(res1$Variances[3], mean(VAR_ratio)), "%", sep = "")

# Example:
# Difference between the two estimations of variance: 
#  - for the number of trees: 2.478042e+12%
#  - for the volume of trees: 2.087256%
#  - for the ratio volume / number of trees: 1559903%
```