### 2024-08-01 CopyOfMC_all.Rmd
MC_all without CI for MC and with coverage rate

---
title: "Monte Carlo and bootstrap for all parameters"
output: pdf_document
date: "2024-07-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of this document is to create a large number of Monte Carlo iterations to produce estimators (for the number of trees, the volume of trees and the ratio volume of trees / number of trees), variances of these estimators and confidence intervals. The same elements are also computed for each bootstrap series derived from a smaller number of Monte Carlo iterations. By comparing the two results, we may ensure the validity of the bootstrap strategy to find the estimation of the variances.

In this document, uniform sampling is used to create a sample and one single circle with a radius of 5 is used to compute local variables.

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column
```

# Preliminaries

```{r}
# Importing the database with the artificial forest
all_trees <- read_excel("~/work/Sim_article2024/Data/artificial_forest_round.xls")
# To get closer to the NFI process, trees whose circumference is under 23.5 cm are not considered
trees <- all_trees[(pi*all_trees$d130 > 23.5),]
```

```{r}
# Functions to produce a local variable
source("~/work/Sim_article2024/Useful_functions/Circle.R")
```

```{r}
# Sample size
n <- 1000
```

```{r}
# True values
cat("Number of trees:", length(trees$x), "\nTotal volume:", sum(trees$v), "\nWhich leads to the following ratio \"Total volume / Number of trees\":", sum(trees$v)/nrow(trees))
```

# Monte Carlo alone

10 000 Monte Carlo iterations are appropriated.

```{r}
### Function to automate Monte Carlo method

MonteCarlo <- function(n, B){
  # n: sample size;
  # B: number of iterations for MC

  p <- sum(rep((1/1000)*(1/1000), n)) # inclusion density

  MC_df <- data.frame(useless = rep(0, n))
  MC_est <- data.frame(useless = rep(0, 3), row.names = c("est_nb", "est_vol", "est_ratio"))

  for (b in 1:B) { # All in one "for" loop
    # Uniform sampling over the whole territory
    a1 <- runif(n, 0, 1000)
    a2 <- runif(n, 0, 1000)
    # Variables of interest derived from the sample
    var <- mapply(circle_all, a1, a2)
    MC_b <- data.frame(x = a1, y = a2, z = t(var)[,1], t = t(var)[,2])
    MC_df <- cbind(MC_df, MC_b)
    colnames(MC_df)[(4*(b-1)+2):(4*(b-1)+2+3)] <- c(paste("x", b, sep = ""), paste("y", b, sep = ""), paste("nb", b, sep = ""), paste("vol", b, sep = ""))
    est_b <- t(data.frame(x = sum(MC_df[4*(b-1)+2+2] /p), y = sum(MC_df[4*(b-1)+2+3] /p), z = (sum(MC_df[4*(b-1)+2+3] /p))/(sum(MC_df[4*(b-1)+2+2]/p))))
    MC_est <- cbind(MC_est, est_b)
    colnames(MC_est)[b+1] <- paste("ech", b, sep = "") 
  }
  MC_df$useless <- NULL
  MC_est$useless <- NULL
  MC_var <- c(var(as.vector(t(MC_est[1,]))), var(as.vector(t(MC_est[2,]))), var(as.vector(t(MC_est[3,]))))

  return(list(`Local Variables` = MC_df, Estimations = MC_est, Variances = MC_var))
  # Local Variables: local variables (number and volume) for each point in each sample
  # Estimations: 3 Estimations (number, volume, ratio volume / number) for each sample
  # Variances: Monte Carlo variance for each of the totals / ratio
}
```

# Bootstrap for each Monte Carlo iteration

1 000 Monte Carlo iterations and 400 bootstrap iterations for each one are appropriated.

```{r}
f_boot <- function(df_sample, B){
  # df_sample: original sample; 
  # B: number of bootstrap samples made thanks to df_sample

  df_boot <- data.frame(useless = rep(0, nrow(df_sample)-1))
  for (b in 1:B) {
    nb <- sample(1:nrow(df_sample), (nrow(df_sample)-1), replace = TRUE)
    # (n-1) points at random
    sample_b <- df_sample[nb,]
    df_boot <- cbind(df_boot, sample_b)
    colnames(df_boot)[((b-1)*4+1 + 1):(b*4+1)] <- c(paste("ech", b, colnames(df_sample)[1], sep = ""), paste("ech", b, colnames(df_sample)[2], sep = ""), paste("ech", b, colnames(df_sample)[3], sep = ""), paste("ech", b, colnames(df_sample)[4], sep = ""))
  }
  df_boot$useless <- NULL
  return(df_boot)
  # Return a data frame with B samples of as many variables as we want (in this application, 4) and of size (n-1)
}
```

```{r}
# Making of estimations for each sample made by the bootstrap method
est_boot <- function(df, p){
  # df: dataframe with 4*B columns and n rows

  n <- nrow(df) + 1 # sample size of the initial sample
  nboot <- length(df)/4 # number of bootstrap samples
  vec_est1 <- c()
  vec_est2 <- c()
  vec_est3 <- c()
  for (c in 1:nboot) {
    vec_est1 <- append(vec_est1, (n/(n-1))*sum(df[,3 + (4*(c-1))]/p))
    vec_est2 <- append(vec_est2, (n/(n-1))*sum(df[,4*c]/p))
    vec_est3 <- append(vec_est3, ((n/(n-1))*sum(df[,4*c]/p))/((n/(n-1))*sum(df[,3 + (4*(c-1))]/p)))
    # (n/(n-1)) is the correction needed since we choose only (n-1) points
  }
  vec_est <- data.frame(est_nb = vec_est1, est_vol = vec_est2, est_ratio = vec_est3)
  return(vec_est)
  # Data frame with 3 estimations for each sample
}
```

```{r}
### Function to automate Monte Carlo + bootstrap method

MC_boot <- function(n, B1, B2){
  # n: sample size;
  # B1: number of iterations for MC;
  # B2: number of iterations for bootstrap

  p <- sum(rep((1/1000)*(1/1000), n))
  BOOT_df <- list() # All local variables
  BOOT_est <- list() # All estimations
  BOOT_var <- list() # Variances
  BOOT_CI <- list() # Confidence Intervals
  for (b in 1:B1) { # One MC iteration at a time
    # Uniform sampling over the whole territory
    a1 <- runif(n, 0, 1000)
    a2 <- runif(n, 0, 1000)
    var <- mapply(circle_all, a1, a2)
    df_b <- data.frame(x = a1, y = a2, z = t(var)[,1], t = t(var)[,2])
    colnames(df_b) <- c(paste("x.MC", b, sep = ""), paste("y.MC", b, sep = ""), paste("nb.MC", b, sep = ""), paste("vol.MC", b, sep = ""))
    est_b <- t(data.frame(x = sum(df_b[3] /p), y = sum(df_b[4] /p), z = (sum(df_b[4] /p))/(sum(df_b[3]/p))))

    # Bootstrap iterations
    df_boot <- f_boot(df_b, B2)
    est_boot_val <- est_boot(df_boot, p)

    BOOT_var[[b]] <- list(var_nb = var(est_boot_val$est_nb), var_vol = var(est_boot_val$est_vol), var_ratio = var(est_boot_val$est_ratio))
    names(BOOT_var)[b] <- paste("Variance", b, sep = "")
    BOOT_df <- c(BOOT_df, list(data_init = df_b, data_boot = df_boot))
    names(BOOT_df)[(2*b-1):(2*b)] <- c(paste("df_MC", b, sep = ""), paste("df_boot", b, sep = ""))
    BOOT_est <- c(BOOT_est, list(est_init_b = est_b, est_boot_b = est_boot_val))
    names(BOOT_est)[(2*b-1):(2*b)] <- c(paste("est_MC", b, sep = ""), paste("est_boot", b, sep = ""))

    # Confidence intervals
    norm_CI <- data.frame(useless = rep(0, 2))
    perc_CI <- data.frame(useless = rep(0, 2))
    rev_perc_CI <- data.frame(useless = rep(0, 2))
    for (l in 1:3) { # Loop: one estimator each time
      sorted_est <- sort(est_boot_val[,l])
      lim_inf <- sorted_est[floor(B2*2.5/100)]
      lim_sup <- sorted_est[B2 - floor(B2*2.5/100)]
      perc_CI <- cbind(perc_CI, data.frame(c(lim_inf, lim_sup)))
      rev_perc_CI <- cbind(rev_perc_CI, data.frame(c(est_b[l,]*2 - lim_sup, est_b[l,]*2 - lim_inf)))
      norm_CI <- cbind(norm_CI, data.frame(c(est_b[l,]-qnorm(0.975)*sqrt(var(est_boot_val[,l])), est_b[l,]+qnorm(0.975)*sqrt(var(est_boot_val[,l])))))
    }
    norm_CI$useless <- NULL
    perc_CI$useless <- NULL
    rev_perc_CI$useless <- NULL
    colnames(norm_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
    colnames(perc_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
    colnames(rev_perc_CI) <- c("lim_nb", "lim_vol", "lim_ratio")
    all_CI <- list(Normal = norm_CI, Percentile = perc_CI, `Reverse percentile` = rev_perc_CI)
    # BOOT_CI <- c(BOOT_CI, CI_b = all_CI)
    BOOT_CI[[b]] <- all_CI
    names(BOOT_CI)[b] <- paste("CI", b, sep = "")
  }
  return(list(`Local Variables` = BOOT_df, Estimations = BOOT_est, Variances = BOOT_var, `Confidence Intervals` = BOOT_CI))
  # Local Variables: local variables (number and volume) for each point in each sample
  # Estimations: 3 estimations (number, volume, ratio volume / number) for each sample
  # Variances: mean of the bootstrap variance for each estimator
  # Confidence Intervals: 3 different forms of 95% confidence intervals, for each total / ratio
}
```

# Example

## Monte Carlo alone

```{r}
start.time <- Sys.time()

res1 <- MonteCarlo(n, 10000)
# Just one circle to begin with

end.time <- Sys.time()
time.taken <- end.time - start.time

time.taken
```

```{r}
cat("Monte Carlo means of estimations: \n - for the number of trees:", mean(as.vector(t(res1$Estimations[1,]))), "\n - for the volume of trees:", mean(as.vector(t(res1$Estimations[2,]))), "\n - for the ratio volume / number of trees:", mean(as.vector(t(res1$Estimations[3,]))))

# Example:
## Monte Carlo means of estimations:
##  - for the number of trees: 29251.69
##  - for the volume of trees: 5407.967
##  - for the ratio volume / number of trees: 0.184873
```

```{r}
cat("Monte Carlo variances: \n - for the number of trees:", res1$Variances[1], "\n - for the volume of trees:", res1$Variances[2], "\n - for the ratio volume / number of trees:", res1$Variances[3])

# Example:
## Monte Carlo variances:
##  - for the number of trees: 2604340
##  - for the volume of trees: 99936.6
##  - for the ratio volume / number of trees: 1.122037e-05
```

## Monte Carlo and bootstrap

```{r}
start.time <- Sys.time()

res2 <- MC_boot(n, B1 = 1000, B2 = 400)
# Just one circle to begin with

end.time <- Sys.time()
time.taken <- end.time - start.time

time.taken
```

```{r}
# We want all the variances obtained for each parameter of interest, 
# to find the mean for each of these series
EST_nb <- c()
EST_vol <- c()
EST_ratio <- c()
for (elt in res2$Estimations) {
  if ("data.frame" %in% class(elt)){ # we target bootstrap samples made from each MC iteration
    EST_nb <- append(EST_nb, mean(elt$est_nb))
    EST_vol <- append(EST_vol, mean(elt$est_vol))
    EST_ratio <- append(EST_ratio, mean(elt$est_ratio))
  }
}
```

```{r}
cat("Mean of the means of estimations obtained through each bootstrap process:\n - for the number of trees:", mean(EST_nb), "\n - for the volume of trees:", mean(EST_vol), "\n - for the ratio volume / number of trees:", mean(EST_ratio))

# Example:
## Mean of the means of estimations obtained through each bootstrap process:
##  - for the number of trees: 29280.52
##  - for the volume of trees: 5411.916
##  - for the ratio volume / number of trees: 0.1848212
```

```{r}
# We want all the variances obtained for each parameter of interest, 
# to find the mean for each of these series
VAR_nb <- c()
VAR_vol <- c()
VAR_ratio <- c()
for (elt in res2$Variances) {
  VAR_nb <- append(VAR_nb, elt$var_nb)
  VAR_vol <- append(VAR_vol, elt$var_vol)
  VAR_ratio <- append(VAR_ratio, elt$var_ratio)
}
```

```{r}
cat("Mean of the variance of bootstrap samples:\n - for the number of trees:", mean(VAR_nb), "\n - for the volume of trees:", mean(VAR_vol), "\n - for the ratio volume / number of trees:", mean(VAR_ratio))

# Example:
## Mean of the variance of bootstrap samples:
##  - for the number of trees: 2562574
##  - for the volume of trees: 98333.62
##  - for the ratio volume / number of trees: 1.121868e-05
```

```{r}
## We want all the CI based on Normal distribution obtained for each parameter of interest, 
# to find the coverage rate
CI_nb <- data.frame(useless = c(0,0))
CI_vol <- data.frame(useless = c(0,0))
CI_ratio <- data.frame(useless = c(0,0))
for (elt in res2$`Confidence Intervals`) {
  CI_nb <- cbind(CI_nb, elt$Normal$lim_nb)
  CI_vol <- cbind(CI_vol, elt$Normal$lim_vol)
  CI_ratio <- cbind(CI_ratio, elt$Normal$lim_ratio)
}
CI_nb$useless <- NULL
CI_vol$useless <- NULL
CI_ratio$useless <- NULL

CI_nb <- as.data.frame(t(CI_nb))
rownames(CI_nb) <- 1:(nrow(CI_nb))
colnames(CI_nb) <- c("x", "y")

CI_vol <- as.data.frame(t(CI_vol))
rownames(CI_vol) <- 1:(nrow(CI_vol))
colnames(CI_vol) <- c("x", "y")

CI_ratio <- as.data.frame(t(CI_ratio))
rownames(CI_ratio) <- 1:(nrow(CI_ratio))
colnames(CI_ratio) <- c("x", "y")
```

```{r}
# Check the coverage rate
TC_nb <- (1/nrow(CI_nb))*sum( mapply(function(a,b){as.numeric((29272 > a) & (29272 < b))}, CI_nb$x, CI_nb$y) )
TC_vol <- (1/nrow(CI_vol))*sum( mapply(function(a,b){as.numeric((5412.722 > a) & (5412.722 < b))}, CI_vol$x, CI_vol$y) )
TC_ratio <- (1/nrow(CI_ratio))*sum( mapply(function(a,b){as.numeric((0.1849112 > a) & (0.1849112 < b))}, CI_ratio$x, CI_ratio$y) )
cat("Concerning the normal approach, the coverage rate are the following: \n - for the number of trees: ", TC_nb, "\n - for the volume of trees: ", TC_vol, "\n - for the ratio volume / number of trees: ", TC_ratio)
```

```{r}
## We want all the Percentile CI obtained for each parameter of interest, 
# to find the coverage rate
CI_perc_nb <- data.frame(useless = c(0,0))
CI_perc_vol <- data.frame(useless = c(0,0))
CI_perc_ratio <- data.frame(useless = c(0,0))
for (elt in res2$`Confidence Intervals`) {
  CI_perc_nb <- cbind(CI_perc_nb, elt$Percentile$lim_nb)
  CI_perc_vol <- cbind(CI_perc_vol, elt$Percentile$lim_vol)
  CI_perc_ratio <- cbind(CI_perc_ratio, elt$Percentile$lim_ratio)
}
CI_perc_nb$useless <- NULL
CI_perc_vol$useless <- NULL
CI_perc_ratio$useless <- NULL

CI_perc_nb <- as.data.frame(t(CI_perc_nb))
rownames(CI_perc_nb) <- 1:(nrow(CI_perc_nb))
colnames(CI_perc_nb) <- c("x", "y")

CI_perc_vol <- as.data.frame(t(CI_perc_vol))
rownames(CI_perc_vol) <- 1:(nrow(CI_perc_vol))
colnames(CI_perc_vol) <- c("x", "y")

CI_perc_ratio <- as.data.frame(t(CI_perc_ratio))
rownames(CI_perc_ratio) <- 1:(nrow(CI_perc_ratio))
colnames(CI_perc_ratio) <- c("x", "y")
```

```{r}
# Check the coverage rate
TC_nb <- (1/nrow(CI_perc_nb))*sum( mapply(function(a,b){as.numeric((29272 > a) & (29272 < b))}, CI_perc_nb$x, CI_perc_nb$y) )
TC_vol <- (1/nrow(CI_perc_vol))*sum( mapply(function(a,b){as.numeric((5412.722 > a) & (5412.722 < b))}, CI_perc_vol$x, CI_perc_vol$y) )
TC_ratio <- (1/nrow(CI_perc_ratio))*sum( mapply(function(a,b){as.numeric((0.1849112 > a) & (0.1849112 < b))}, CI_perc_ratio$x, CI_perc_ratio$y) )
cat("Concerning the percentile approach, the coverage rate are the following: \n - for the number of trees: ", TC_nb, "\n - for the volume of trees: ", TC_vol, "\n - for the ratio volume / number of trees: ", TC_ratio)
```

```{r}
## We want all the Reverse Percentile CI obtained for each parameter of interest, 
# to find the coverage rate
CI_nb <- data.frame(useless = c(0,0))
CI_vol <- data.frame(useless = c(0,0))
CI_ratio <- data.frame(useless = c(0,0))
for (elt in res2$`Confidence Intervals`) {
  CI_nb <- cbind(CI_nb, elt$`Reverse percentile`$lim_nb)
  CI_vol <- cbind(CI_vol, elt$`Reverse percentile`$lim_vol)
  CI_ratio <- cbind(CI_ratio, elt$`Reverse percentile`$lim_ratio)
}
CI_nb$useless <- NULL
CI_vol$useless <- NULL
CI_ratio$useless <- NULL

CI_nb <- as.data.frame(t(CI_nb))
rownames(CI_nb) <- 1:(nrow(CI_nb))
colnames(CI_nb) <- c("x", "y")

CI_vol <- as.data.frame(t(CI_vol))
rownames(CI_vol) <- 1:(nrow(CI_vol))
colnames(CI_vol) <- c("x", "y")

CI_ratio <- as.data.frame(t(CI_ratio))
rownames(CI_ratio) <- 1:(nrow(CI_ratio))
colnames(CI_ratio) <- c("x", "y")
```

```{r}
# Check the coverage rate
TC_nb <- (1/nrow(CI_nb))*sum( mapply(function(a,b){as.numeric((29272 > a) & (29272 < b))}, CI_nb$x, CI_nb$y) )
TC_vol <- (1/nrow(CI_vol))*sum( mapply(function(a,b){as.numeric((5412.722 > a) & (5412.722 < b))}, CI_vol$x, CI_vol$y) )
TC_ratio <- (1/nrow(CI_ratio))*sum( mapply(function(a,b){as.numeric((0.1849112 > a) & (0.1849112 < b))}, CI_ratio$x, CI_ratio$y) )
cat("Concerning the reverse percentile approach, the coverage rate are the following: \n - for the number of trees: ", TC_nb, "\n - for the volume of trees: ", TC_vol, "\n - for the ratio volume / number of trees: ", TC_ratio)
```

## Comparison of the variances

```{r}
diff_percent <- function(x,y)100*(max(x,y)-min(x,y))/min(x,y)
cat("Difference between the two estimations of variance: \n - for the number of trees: ", diff_percent(res1$Variances[1], mean(VAR_nb)), "%", "\n - for the volume of trees: ", diff_percent(res1$Variances[2], mean(VAR_vol)), "%\n - for the ratio volume / number of trees: ", diff_percent(res1$Variances[3], mean(VAR_ratio)), "%", sep = "")

# Example:
## Difference between the two estimations of variance:
##  - for the number of trees: 1.629866%
##  - for the volume of trees: 1.630143%
##  - for the ratio volume / number of trees: 0.01505727%
```