### 2024-09-12 Intermediary_steps/Comparison_test.Rmd

---
title: "Comparison_test"
output: pdf_document
date: "2024-08-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column
```

# Preliminaries

```{r}
# Importing the database with the artificial forest
all_trees <- read_excel("~/work/Sim_article2024/Data/artificial_forest_round.xls")
# To get closer to the NFI process, trees whose circumference is under 23.5 cm are not considered
trees <- all_trees[(pi*all_trees$d130 > 23.5),]
```

```{r}
# Functions to produce a local variable
source("~/work/Sim_article2024/Useful_functions/Circle.R")
```

```{r}
# Sample size
n <- 1000
a1 <- runif(n, 0, 1000)
a2 <- runif(n, 0, 1000)
```

# Comparison

```{r}
start.time <- Sys.time()

var1 <- mapply(circle_all, a1, a2)

end.time <- Sys.time()
time.taken <- end.time - start.time

time.taken
```

```{r}
start.time <- Sys.time()

var3 <- mapply(circle_ALL_bis, a1, a2)

end.time <- Sys.time()
time.taken <- end.time - start.time

time.taken
```

```{r}
start.time <- Sys.time()

var5 <- mapply(circle_ALL_quater, a1, a2)

end.time <- Sys.time()
time.taken <- end.time - start.time

time.taken
```

```{r}
start.time <- Sys.time()

var2 <- mapply(circle_ALL, a1, a2)

end.time <- Sys.time()
time.taken <- end.time - start.time

time.taken
```