### 2024-09-12 Intermediary_steps/Ratio_boot.Rmd

---
title: "Bootstrap method with estimated ratios"
output: pdf_document
date: "2024-07-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column
```

# Preliminaries

```{r}
# Importing the database with the artificial forest
trees <- read_excel("~/work/Sim_article2024/Data/artificial_forest_round.xls")
```

```{r}
# Functions to produce a local variable
source("~/work/Sim_article2024/Useful_functions/Circle.R")
```

```{r}
# Sample size
n <- 1000

# Inclusion density
p <- sum(rep((1/1000)*(1/1000), n))
```

# Ratio volume/number of trees

```{r}
# True ratio
cat("True ratio \"total volume over number of trees\":", sum(trees$v)/nrow(trees))
```

```{r}
random_points <- data.frame(x = runif(n, 0, 1000), y = runif(n, 0, 1000))
```

```{r}
# An estimation of the ratio
a1 <- mapply(circle_nb, random_points[, 1], random_points[, 2])
a1 <- sum(a1/p)
a2 <- mapply(circle_vol, random_points[, 1], random_points[, 2])
a2 <- sum(a2/p)
est_ratio <- a2/a1
# 0.1761545 e.g.
```

```{r}
# A second estimation of the ratio with 3 categories
b1 <- mapply(circle_nb_3cats, random_points[, 1], random_points[, 2])
b11 <- sum(b1[1,]/p)
b12 <- sum(b1[2,]/p)
b13 <- sum(b1[3,]/p)
b2 <- mapply(circle_vol_3cats, random_points[, 1], random_points[, 2])
b21 <- sum(b2[1,]/p)
b22 <- sum(b2[2,]/p)
b23 <- sum(b2[3,]/p)

est_vol <- b11*b21/a1 + b12*b22/a1 + b13*b23/a1
```

```{r}
# Underlying model?
reg <- lm(v ~ d130, data = trees)
cat("R^2:", summary(reg)$r.squared)
```